---
title: "data clean"
output: html_notebook
---

### libraries

```{r}
library(tidyverse)
```

## Trust

### Trust news in social media most of the time

**Scale:**

+ Strongly disagree

+ Disagree

+ Tend to disagree

+ Neither agree nor disagree

+ Tend to agree

+ Agree

+ Strongly agree

```{r}
social_trust <- read_csv("data/trust/social-news-trust.csv")

# stack gender
social_trust_gender <- social_trust %>% select(question,question_choice, male, female) %>% gather(gender, count, male, female)

# stack age
social_trust_age <- social_trust %>% select(question, question_choice, `18-24`:`65+`) %>% gather(age, count, `18-24`:`65+`)

write_csv(social_trust_age, "data/trust/age-socialtrust-split.csv")

```

### Trust news in search engines most of the time

**Scale:**

+ Strongly disagree

+ Disagree

+ Tend to disagree

+ Neither agree nor disagree

+ Tend to agree

+ Agree

+ Strongly agree

```{r}
search_trust <- read_csv("data/trust/search-news-trust.csv")

# stack gender
search_trust_gender <- search_trust %>% select(question, question_choice, male, female) %>% gather(gender, count, male, female)

# stack age
search_trust_age <- search_trust %>% select(question, question_choice, `18-24`:`65+`) %>% gather(age, count, `18-24`:`65+`)

write_csv(search_trust_age, "data/trust/age-searchtrust-split.csv")

```

### Trust(6-10) vs. Distrust(0-4) in brands

**Scale:** 1-10 

**Options:**

+ Fox News

+ CNN

+ NBC/MSNBC News

+ ABC News

+ CBS News

+ Yahoo! News

+ Local television news

+ New York Times

+ Washington Post

+ HuffPost (Huffington Post)

+ Buzzfeed News

+ NPR News

+ Breitbart

+ Wall Street Journal

+ Vice News

```{r}
brand_trust <- read_csv("data/trust/brand-trust-distrust.csv")

# stack gender
brand_trust_gender <- brand_trust %>% select(question, question_choice, male, female) %>% gather(gender, count, male, female)

# stack age
brand_trust_age <- brand_trust %>% select(question, question_choice, `18-24`:`65+`) %>% gather(age, count, `18-24`:`65+`)


```

-----

## Misinformation

**Agree with statements:**

+ Technology companies (e.g. Facebook and Google) should do more to make it easier to separate what is real and what is fake on the internet

+ Media companies and journalists should do more to make it easier to separate what is real and what is fake on the internet

+ The Government should do more to make it easier to separate what is real and fake on the internet

**Scale:**

+ Strongly disagree

+ Disagree

+ Tend to disagree

+ Neither agree nor disagree

+ Tend to agree

+ Agree

+ Strongly agree

```{r}
misinfo_trio <- read_csv("data/misinfo/tech-media-gov.csv")

# stack gender
misinfo_gender <- misinfo_trio %>% select(entity, question_choice, Male, Female) %>% gather(gender, count, Male, Female)

# stack age
misinfo_age <- misinfo_trio %>% select(entity, question_choice, `18-24`:`65+`) %>% gather(age, count, `18-24`:`65+`)

# separate by age groups

misinfo_age_tech <- misinfo_age %>% filter(entity == "Technology companies")
write_csv(misinfo_age_tech, "data/misinfo/trio-tech.csv")

misinfo_age_media <- misinfo_age %>% filter(entity == "Media companies and journalists")
write_csv(misinfo_age_media, "data/misinfo/trio-media.csv")

misinfo_age_gov <- misinfo_age %>% filter(entity == "Government")
write_csv(misinfo_age_gov, "data/misinfo/trio-gov.csv")
```

-----

## Literacy
